---
title: "HW04"
author: "Elyse Adamic"
date: "03/10/2019"
output: 
  html_document:
    keep_md: true
    toc: true
---
```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(gapminder)
```

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

### Exercise 1: Univariate Pivoting
1) Put data in wider format.
2) Produce a plot of the wide data.
3) Re-lengthen the wider data. 

Option 1: 
Make a tibble with one row per year, and columns for life expectancy for two or more countries. Note when we put the data into wider format, we lose the other variables in the dataset so the select isn't actually needed. Also we have no way to recover these when pivoting back to long format. 

```{r}
(wide_uni <- gapminder %>% 
  filter(country == "Canada"| country ==  "Switzerland"| country == "United States") %>% 
  select(country, year, lifeExp) %>% 
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = lifeExp))

wide_uni %>% 
  ggplot(aes(Switzerland,Canada)) +
    geom_point(shape = 21, fill = "red", alpha = 0.5)

wide_uni %>% 
  pivot_longer(cols = c("Canada", "Switzerland", "United States" ),
               names_to = "country",
               values_to = "LifeExp") %>% 
  arrange(country, year)
```

### Exercise 2: Multivariate Pivoting
1) Put data in wider format.
3) Re-lengthen the wider data. 

Option 1: Make a tibble with one row per year, and columns for life expectancy and GDP per capita (or two other numeric variables) for two or more countries. Note here we want to "retain column status" so we use the ".value" and we do not need a values_to. 

```{r}
(wide_multi <- gapminder %>% 
  filter(country == "Canada"| country ==  "Switzerland") %>% 
  pivot_wider(id_cols = year,
              names_from = country,
              names_sep   = "_",
              values_from = c(lifeExp, gdpPercap)))

(wide_multi %>% 
  pivot_longer(cols = c(-year),
               names_to = c(".value","country"),
               names_sep = "_",) %>% 
  arrange(country, year))
```

### Exercise 3: Table Joins

```{r, include = FALSE}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")

```

#### 3.1
For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble.

- this will be a mutating join since we want a new variable 
- this will be a left outer join and guestlist is our "main" tibble

We first had to separate email_name and rename guest so there was a common variable to join by, note also the space after the comma was needed here for separating, or else names woudl not match. Another way would be to define an index for party in email 

```{r}
email_name <- email %>% 
  separate_rows(guest, sep = ", ") %>% 
  rename(name = guest)

guest %>% 
  left_join(email_name,
            by = "name")
```

#### 3.2
Who do we have emails for, yet are not on the guestlist?
- start with email, antijoin guest

```{r}
email_name %>% 
  anti_join(guest,
            by = "name")

```

#### 3.3
Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist).

- check: this should be 33 rows long

```{r}
guest %>% 
  full_join(email_name,
            by = "name") %>% 
  DT::datatable()
```
